---
title: "Hw03-hyeongcheol-park"
author: "tom"
date: '2017-09-30'
output: html_document
---
####warming up- install packages
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```



### 1.Let's get the maximum and minimum of GDP per capita for all continents.

```{r}
(q1<-gapminder %>% 
  group_by(continent, year) %>% 
  mutate(Maxgdp=max(gdpPercap),Mingdp=min(gdpPercap)) %>% 
  select(continent, year, Maxgdp, Mingdp))
```
Max and Min of GDP/capita for each continent per each year!

I will make a graph.

```{r}
q1 %>% 
  ggplot(aes(color=continent))+facet_wrap(~ continent) +geom_point(aes(year,Maxgdp))+geom_point(aes(year,Mingdp))+scale_y_continuous("Max and Min")
```
  <br />We can easily see the difference between Min and Max of GDP/capita on each continent per year.

 But it seems messy. so I will pick only in 20C and 20C and make a plot.
 
```{r}
q1 %>% 
  filter(year %in% 1992:2017) %>% 
  ggplot(aes(color=continent, group=continent)) +geom_point(aes(year,Maxgdp))+geom_point(aes(year,Mingdp))+scale_y_continuous("Max and Min")
```
 

### 2.Look at the spread of GDP per capita within the continents.

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(mean_gdp=mean(gdpPercap), sd_gdp=sd(gdpPercap))
```
It`s a brief summary for GDP/capita within the continents.





```{r}
gapminder %>% 
  ggplot(aes(year, gdpPercap,group=country,color=continent))+facet_wrap( ~ continent) +geom_point()+geom_line()
```
 <br /> It shows the spread of GDP per capita within the continents
```{r}
ggplot(gapminder, aes(year, gdpPercap,
                      colour=continent,group=country)) +
    geom_point()+geom_line(alpha=0.2)
```
  <br />Or we can use this.


```{r}
gapminder %>%
    ggplot(aes(x = continent, y = gdpPercap)) + geom_boxplot() + 
  geom_jitter(position = position_jitter(width = 0.2, height = 0), alpha = 0.3, size = 2.5, colour = "green") +
  labs(title = "GDP Per Capita vs Continent", x = "Continent", y = "GDP Per capita" )
```
 <br /> Maybe boxplot is better choice.
 
```{r}
gapminder %>%
  ggplot(aes(x = gdpPercap, fill = continent)) +
  geom_density (alpha = 0.4) +
  labs (title = "GDP/capita for each continent", x = "GDP/capita", y = "Density")
```
 
 
### 3.Compute a trimmed mean of life expectancy for different years. 
###How is life expectancy changing over time on different continents?
```{r}
(newlife<-gapminder %>% 
  group_by(country) %>% 
 mutate(newlife_exp=lifeExp*year[year==2007]/year))
```
 <br /> As time goes by, people tend to live longer. So what I did is lifeExp*year[year==2007]/year, which will correct the tendency a little bit.
(Actually I did it by using gdp, but it scales lifeExp too much and I got data on which people at 1952 live longer than people in 2007! so it seems not proper way.)

```{r}
newlife %>% 
  ggplot(aes(year,newlife_exp,color=gdpPercap,group=country))+facet_wrap(~ continent)+geom_point()+geom_line()
```

  <br />Still there is the tendency that comtemporary people live longer.






### 4.Report the absolute and/or relative abundance of countries with low life expectancy over time by continent

**I will use median lifeExp for each continent.**



```{r}
(belowage<-newlife %>% 
  group_by(continent) %>% 
  mutate(meanage=median(lifeExp)))
```
  <br />I got the average of lifeExp for every continent.

```{r}
(below<-belowage %>% 
  filter(newlife_exp <= meanage))
```

<br />there are 792 countries whose lifeExp is below their continent's median, from 1952 to 2007.


```{r}
below %>% 
  group_by(year, continent) %>% 
  summarize(num=n())
```

<br />I got for each year, how many countries has lower lifeExp than median lifeExp for their continent!

```{r}

knitr::kable(below %>% 
  group_by(year, continent) %>% 
  summarize(num=n()))
```




<br />pretty chart.And for graph,

```{r}
below %>% 
  group_by(year, continent) %>% 
  summarize(num=n()) %>% 
  ggplot(aes(year,num, colour=continent,group=continent))+geom_point()+geom_line()+labs(y="number of countries", title="how many countries people die early")
```
<br /> Early dying people seem to be getting smaller. 



## 5.Find countries with interesting stories. Open-ended and, therefore, hard. Promising but unsuccessful attempts are encouraged. This will generate interesting questions to follow up on in class.

Make up your own! Between the dplyr coverage in class and the list above, I think you get the idea.




Notice which figures are easy/hard to make, which data formats make better inputs for plotting functions vs. for human-friendly tables.

But I want to do more!
Layout stretch goal: get table and figure side-by-side. This gist might get you started.

Table stretch goal: there are some really nice fancy table helper packages. This tweet from @polesasunder will point you toward some R packages you may want to check out (pander, xtable, stargazer).

Report your process
You’re encouraged to reflect on what was hard/easy, problems you solved, helpful tutorials you read, etc. Give credit to your sources, whether it’s a blog post, a fellow student, an online tutorial, etc.


As before, make a new issue and call it "Homework 03 is ready for grading".
