---
title: "Hw03-hyeongcheol-park"
author: "tom"
date: '2017-09-30'
output: html_document
---
####warming up- install packages
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```



### 1.Let's get the maximum and minimum of GDP per capita for all continents.

```{r}
(q1<-gapminder %>% 
  group_by(continent, year) %>% 
  mutate(Maxgdp=max(gdpPercap),Mingdp=min(gdpPercap)) %>% 
  select(continent, year, Maxgdp, Mingdp))
```
Max and Min of GDP/capita for each continent per each year!

I will make a graph.

```{r}
q1 %>% 
  ggplot(aes(color=continent))+facet_wrap(~ continent) +geom_point(aes(year,Maxgdp))+geom_point(aes(year,Mingdp))+scale_y_continuous("Max and Min")
```
  <br />We can easily see the difference between Min and Max of GDP/capita on each continent per year.

 But it seems messy. so I will pick only in 20C and 20C and make a plot.
 
```{r}
q1 %>% 
  filter(year %in% 1992:2017) %>% 
  ggplot(aes(color=continent, group=continent)) +geom_point(aes(year,Maxgdp))+geom_point(aes(year,Mingdp))+scale_y_continuous("Max and Min")
```
 

### 2.Look at the spread of GDP per capita within the continents.

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(mean_gdp=mean(gdpPercap), sd_gdp=sd(gdpPercap), min_gdp=min(gdpPercap),max_gdp=max(gdpPercap),median_gdp=median(gdpPercap))
```
It`s a brief summary for GDP/capita within the continents.





```{r}
gapminder %>% 
  ggplot(aes(year, gdpPercap,group=country,color=continent))+facet_wrap( ~ continent) +geom_point()+geom_line()
```
 <br /> It shows the spread of GDP per capita within the continents
```{r}
ggplot(gapminder, aes(year, gdpPercap,
                      colour=continent,group=country)) +
    geom_point()+geom_line(alpha=0.2)
```
  <br />Or we can use this.


```{r}
gapminder %>%
    ggplot(aes(x = continent, y = gdpPercap)) + geom_boxplot() + 
  geom_jitter(position = position_jitter(width = 0.2, height = 0), alpha = 0.3, size = 2.5, colour = "green") +
  labs(title = "GDP Per Capita vs Continent", x = "Continent", y = "GDP Per capita" )
```
 <br /> Maybe boxplot is better choice.
 
```{r}
gapminder %>%
  ggplot(aes(x = gdpPercap, fill = continent)) +
  geom_density (alpha = 0.4) +
  labs (title = "GDP/capita for each continent", x = "GDP/capita", y = "Density")
```
  <br /> I am trying many things that I have learned from peer reviewing.
 
### 3.Compute a trimmed mean of life expectancy for different years. 
###How is life expectancy changing over time on different continents?
```{r}
(newlife<-gapminder %>% 
  group_by(country) %>% 
 mutate(newlife_exp=lifeExp*year[year==2007]/year))
```
 <br /> As time goes by, people tend to live longer. So what I did is lifeExp*year[year==2007]/year, which will correct the tendency a little bit.
(Actually I did it by using gdp, but it scales lifeExp too much and I got data on which people at 1952 live longer than people in 2007! so it seems not proper way.)

```{r}
newlife %>% 
  ggplot(aes(year,newlife_exp,color=gdpPercap,group=country))+facet_wrap(~ continent)+geom_point()+geom_line()
```

  <br />Still there is the tendency that comtemporary people live longer.






### 4.Report the absolute and/or relative abundance of countries with low life expectancy over time by continent

**I will use median lifeExp for each continent.**



```{r}
(belowage<-newlife %>% 
  group_by(continent) %>% 
  mutate(meanage=median(lifeExp)))
```
  <br />I got the average of lifeExp for every continent.

```{r}
(below<-belowage %>% 
  filter(newlife_exp <= meanage))
```

<br />there are 792 countries whose lifeExp is below their continent's median, from 1952 to 2007.


```{r}
below %>% 
  group_by(year, continent) %>% 
  summarize(num=n())
```

<br />I got for each year, how many countries has lower lifeExp than median lifeExp for their continent!

```{r}

knitr::kable(below %>% 
  group_by(year, continent) %>% 
  summarize(num=n()))
```




<br />pretty chart.And for graph,

```{r}
below %>% 
  group_by(year, continent) %>% 
  summarize(num=n()) %>% 
  ggplot(aes(year,num, colour=continent,group=continent))+geom_point()+geom_line()+labs(y="number of countries", title="how many countries people die early")
```
<br /> Early dying people seem to be getting smaller. 



## 5.Find countries with interesting stories.
```{r}
gapminder %>% 
  mutate(diff=gdpPercap-lifeExp*10^2) %>% 
  group_by(year) %>% 
  summarize(max(diff))
  
  
```
I got Max GDP+ Min LifeExp from each country.



##6.Make up your own! Between the dplyr coverage in class and the list above, I think you get the idea.


##To report my process.

 The most difficult thing was that I would like to make a chart of "how many countries people die early" without overlapping same years five times for each continent. I tried to do it but I gave up to find the solution. 
  And It seems that I am getting use to using %>% function. It was not fully free for me to analize whatever I want to try, but I think I am getting better for that.
  However, honestly, I still don't know well how to decorate plots  changing axis, scale, and so on. So I skipped many of that in my assignment3.
  Helpful tutorial was absolutely the lecture note, [here](https://stat545.com/cm007-notes_and_exercises.html). And I searched some [blog](http://www.dodomira.com/2016/03/18/ggplot2-기초/), which was more readable for me, while I don't think it will be helpful to others.


