---
title: "Hw03-hyeongcheol-park"
author: "tom"
date: '2017-09-30'
output: github_document
---
####warming up- install packages
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```



### 1.Let's get the maximum and minimum of GDP per capita for all continents!

```{r}
(q1<-gapminder %>% 
  group_by(continent, year) %>% 
  mutate(Maxgdp=max(gdpPercap),Mingdp=min(gdpPercap)) %>% 
  select(continent, year, Maxgdp, Mingdp))
```
Max and Min of GDP/capita for each continent per each year!

I will make a graph.

```{r}
q1 %>% 
  ggplot(aes(color=continent))+facet_wrap(~ continent) +geom_point(aes(year,Maxgdp))+geom_point(aes(year,Mingdp))+scale_y_continuous("Max and Min")
```
  <br />We can easily see the difference between Min and Max of GDP/capita on each continent per year.

 But it seems messy. so I will pick only in 20C and 20C and make a plot.
 
```{r}
q1 %>% 
  filter(year %in% 1992:2017) %>% 
  ggplot(aes(color=continent, group=continent)) +geom_point(aes(year,Maxgdp))+geom_point(aes(year,Mingdp))+scale_y_continuous("Max and Min")
```
 

### 2.Look at the spread of GDP per capita within the continents.

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(mean_gdp=mean(gdpPercap), sd_gdp=sd(gdpPercap), min_gdp=min(gdpPercap),max_gdp=max(gdpPercap),median_gdp=median(gdpPercap))
```
It`s a brief summary for GDP/capita within the continents.





```{r}
gapminder %>% 
  ggplot(aes(year, gdpPercap,group=country,color=continent))+facet_wrap( ~ continent) +geom_point()+geom_line()
```
 <br /> It shows the spread of GDP per capita within the continents
```{r}
ggplot(gapminder, aes(year, gdpPercap,
                      colour=continent,group=country)) +
    geom_point()+geom_line(alpha=0.2)
```
  <br />Or we can use this.


```{r}
gapminder %>%
    ggplot(aes(x = continent, y = gdpPercap)) + geom_boxplot() + 
  geom_jitter(position = position_jitter(width = 0.2, height = 0), alpha = 0.3, size = 2.5, colour = "green") +
  labs(title = "GDP Per Capita vs Continent", x = "Continent", y = "GDP Per capita" )
```
 <br /> Maybe boxplot is better choice.
 
```{r}
gapminder %>%
  ggplot(aes(x = gdpPercap, fill = continent)) +
  geom_density (alpha = 0.4) +
  labs (title = "GDP/capita for each continent", x = "GDP/capita", y = "Density")
```
  <br /> I am trying many things that I have learned from peer reviewing.
 
### 3.Compute a trimmed mean of life expectancy for different years. 
###How is life expectancy changing over time on different continents?
```{r}
(newlife<-gapminder %>% 
  group_by(country) %>% 
 mutate(newlife_exp=lifeExp*year[year==2007]/year))
```
 <br /> As time goes by, people tend to live longer. So what I did is lifeExp*year[year==2007]/year, which will correct the tendency a little bit.
(Actually I did it by using gdp, but it scales lifeExp too much and I got data on which people at 1952 live longer than people in 2007! so it seems not proper way.)

```{r}
newlife %>% 
  ggplot(aes(year,newlife_exp,color=gdpPercap,group=country))+facet_wrap(~ continent)+geom_point()+geom_line()
```

  <br />Still there is the tendency that comtemporary people live longer.






### 4.Report the absolute and/or relative abundance of countries with low life expectancy over time by continent

**I will use median lifeExp for each continent.**



```{r}
(belowage<-newlife %>% 
  group_by(continent) %>% 
  mutate(meanage=median(lifeExp)))
```
  <br />I got the average of lifeExp for every continent.

```{r}
(below<-belowage %>% 
  filter(newlife_exp <= meanage))
```

<br />there are 792 countries whose lifeExp is below their continent's median, from 1952 to 2007.


```{r}
below %>% 
  group_by(year, continent) %>% 
  summarize(num=n())
```

<br />I got for each year, how many countries has lower lifeExp than median lifeExp for their continent!

```{r}

knitr::kable(below %>% 
  group_by(year, continent) %>% 
  summarize(num=n()))
```




<br />pretty chart.And for graph,

```{r}
below %>% 
  group_by(year, continent) %>% 
  summarize(num=n()) %>% 
  ggplot(aes(year,num, colour=continent,group=continent))+geom_point()+geom_line()+labs(y="number of countries", title="how many countries people die early")
```
<br /> Early dying people seem to be getting smaller. 



## 5.Find countries with interesting stories.
```{r}
gapminder %>% 
  mutate(diff=gdpPercap-lifeExp*10^2) %>% 
  arrange(desc(diff)) %>% 
  select(year,country,diff,lifeExp,gdpPercap)
  
  
```
I made some well-being index, diff, which means higher gdp with longer lifeExp. That is to say, some countries where people should be happy in common sense.

 It turns out Kuwait has many ranks, so I made a plot only for Kuwait.
```{r}
gapminder %>% 
  mutate(diff=gdpPercap-lifeExp*10^2) %>% 
  arrange(desc(diff)) %>% 
  filter(country=="Kuwait") %>% 
  ggplot() + geom_point(aes(year,gdpPercap),alpha = 0.25)+
  geom_point(aes(year,lifeExp*10^3))+
    labs(title="The situation of Kuwait\nblack=lifeExp\ngrey=GDP") + scale_y_continuous("GDP and lifeExp")+
    theme(axis.text.x = element_text(angle=90),
          plot.title = element_text(hjust=0.5),
          legend.position = "bottom")
```
<br /> So the Black colour is for lifeExp, scaled 10^3, and grey for Gdp.
 



##To report my process.

 The most difficult thing was that I would like to make a chart of "how many countries people die early" without overlapping same years five times for each continent. I tried to do it but I gave up to find the solution. 
  And It seems that I am getting use to using %>% function. It was not fully free for me to analize whatever I want to try, but I think I am getting better for that.
  However, honestly, I still don't know well how to decorate plots  changing axis, scale, and so on. So I skipped many of that in my assignment3.
  Helpful tutorial was absolutely the lecture note, [here](https://stat545.com/cm007-notes_and_exercises.html). And I searched some [blog](http://www.dodomira.com/2016/03/18/ggplot2-기초/), which was more readable for me, while I don't think it will be helpful to others. Moreover, my peer review was helpful! here it comes HeatherSummer's homework2 [HeatherSummers](https://github.com/heathersummers/STAT545-hw-Summers-Heather/blob/master/hw02/hw02.md)


